---
title: "XX#####_Sample_Mapping"
author: "curator_name"
date: "`r format(Sys.time(), '%Y-%B-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# General setup

### Updates
**
Note to curators:  
* Describe the updates to the study.  
* Suggested format: date as YYYY-MM-DD and bullet points of changes in this update.  
**

### Summary

1. Study XX##### for [Theme Molecule] - Phase XXX ([study path on fileshare]) has data for the following assays: [names of assay folders, e.g. Cobas, Facs, Fluidigm, Histology, Nanostring, Massspec, Immunoassay, NGS etc].  
[(Sr.) BOM/BOPM/...]: [...]  
Study title: [...]  
  
2. Workflows:  

a. [...] is the central lab for tissue samples. [...] is the central lab for blood samples.  
b. Histogenex performed histology using tissue samples received from clinical sites.  
c. EA RNAseq: Histogenex extracted RNA from tissue samples and sent RNA to EA (now Q2SO) for RNAseq.  
d. EA WES: Histogenex extracted DNA from tissue samples and sent DNA to EA for WES.  
e. FMI: Histogenex extracted DNA from tissue samples and sent DNA to FDMI.  
f. Nanostring: Precision extracted RNA from PBMC samples and sent RNA sampless to LabCorp for Nanostring

3. Decisions made in the sample mapping process  
**  
Notes to curators:  
* Note anything "unusual" encountered during the sample mapping process
**  

### R markdown setup and package loading

```{r setup, include=TRUE}


  library(readr)
  library(readxl)
  library(dplyr)
  library(stringr)
  library(devtools)
  library(tidyr)
  library(haven)
  library(kableExtra)
  library(knitr)


```


```{r functions}

# Create a new binary pipe operator
`%notin%` = Negate(`%in%`)

# str_ignore function - opposite of str_subset - excludes elements of a string vector that have specific pattern
str_ignore <- function(string, pattern) {
  string[!str_detect(string, pattern)]
}

# Date and time:
bsed_format <- "%Y-%m-%d"
bsed_pattern <- "[0-9]{4}-[0-1][0-9]-[0-9]{2}"
bsetm_format <- "%H:%M"
bsetm_pattern <- "[0-9]{2}:[0-9]{2}"

sample_mapping_table_colnames <- c()

# Functions for stitching together data files. Pass a list of files and it outputs a dataframe containing all the data from the files.
# csv
CombineCSVFiles <- function(file_list){
  df <-  data.frame(stringsAsFactors=FALSE)
  for(i in file_list){
    j = as.data.frame(read_csv(i, col_types = cols(.default = col_character())))
    df <- bind_rows(df, j) # bind_rows can bind data frames with mismatched numbers of columns
    #df$filename <- i
  }
  return (df)
}

# txt
CombineTXTFiles <- function(file_list){
  df <-  data.frame(stringsAsFactors=FALSE)
  for(i in file_list){
    j = as.data.frame(read_tsv(i, col_types = cols(.default = col_character())))
    # bind_rows can bind data frames with mismatched numbers of columns
    df <- bind_rows(df, j)
  }
  return (df)
}

# xlsx
CombineXLSXFiles <- function(file_list, sheets, skip = 0){
  df <-  data.frame(stringsAsFactors=FALSE)
  if(missing(sheets)){
    for(i in file_list){
      for(sheet in excel_sheets(i)){
        j <- as.data.frame(read_excel(i, sheet = sheet, col_types = "text", skip = skip))
        # bind_rows can bind data frames with mismatched numbers of columns
        df <- bind_rows(df, j)
      }
    }
  } else {
    for(i in file_list){
      for(sheet in sheets){
        j <- as.data.frame(read_excel(i, sheet = sheet, col_types = "text", skip = skip))
        # bind_rows can bind data frames with mismatched numbers of columns
        df <- bind_rows(df, j)
        }
      }
    }
  return (df)
}

# Function to render a horizontally and vertically scrollable table in knitted html
render_table <- function(df){
  kbl(df) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = T,
                fixed_thead = T, font_size = 7, ) %>%
  scroll_box(width = "100%", height = "200px")
}

# Function to compare output tables. Pass 2 tables you need compared and a key to use for comparison (optional) and it will print the values that differ between the 2 tables.
# Example of usage 1:
# compare_tables(output_table, old_dsi)
# Example of usage 2:
# compare_tables(qlab_hist_tracker_df, ihc_hist_qlab_df, key = c("HIST_ID"))
# Use this line with this function if you'd like to have all lines printed: 
# options(tibble.print_max = Inf, tibble.width = Inf)

compare_tables <- function(df1, df2, key = c("SAMPLE_ID")){
    # Changed metadata
    df12_join <- df1 %>% inner_join(df2, by = key)
    
    for (clmn in names(df12_join)[grepl("\\.x", names(df12_join))]){
      # check which samples have values differ between two tables
      print(str_remove(clmn, ".x"))
      clmn1 <- str_replace(clmn, ".x", ".y")
      rows_that_differ <- which(df12_join[, clmn] != df12_join[, clmn1])
      rows_that_differ_no <- length(rows_that_differ)
      print(rows_that_differ_no)
      if(rows_that_differ_no != 0){
        df <- df12_join[rows_that_differ, c(key,clmn,clmn1)]
        print(df)
        }
    }
}

# Function that selects and adds only those columns that are missing.
# Added columns are empty, don't have any values.
# Example of usage:
# add_columns(output_sample_mapping_table_colnames)
 
add_columns <- function(data, cname){
  add <- cname[!cname %in% names(data)]
  
  if(length(add)!=0) data[add] <- NA
  data
}

# Function to select files dated before the cutoff date
select_files_by_date <- function(files, date){
  files_info <- file.info(files)
  # select files before the given date
  selected_files <- row.names(files_info)[as.Date(files_info$ctime) < date] 
  files <- files[files %in% selected_files]
  return(files)
}

```

### File sources

**
Notes to curators:  
* List all the used and checked resources (even those where no files of interest were found).  

Helpful to include to facilitate QC and to keep for reference:  
* Source of the file, file name and a full path to the file (if available).
* Date the file was downloaded. Nothing was downloaded? Put the date the source was checked.
* Was the source useful? Options: Used/Not used/Not available/Not requested
* Other helpful details: information from the file that was used, the reason why the file was not used. 
See the reference list below.  
**

### File paths and variables

```{r file paths}

cutoff_date <- as.Date("YYYY-MM-DD") # Time stamp for the update.Files added to the study folder after the cut off date are ignored in the current update.
study <- tolower("XX#####")

### input files
dm_filepath <- file.path('C:\\Projects\\Project_2\\dm')
tv_filepath <- file.path('C:\\Projects\\Project_2\\tv')
zb_filepath <- file.path('C:\\Projects\\Project_2\\zb')
tdq_path <- file.path('C:\\Projects\\Project_2\\tdq')
cova_inventory_filepath <- file.path('C:\\Projects\\Project_2\\cova')
hist_rawdata_path
ihc_qlab_folders
ngs_rawdata_path
qlab_hist_tracker_filepath
ngs_rawdata_path


### output files

# output_table_path <- paste0('C:\\Projects\\Project_2', 'result.csv')      

```

# Data wrangling

## SDTMv

### Subjects
```{r}

# Collect USUBJIDs and extract PATNUMs. Add SCRNUMs in next section from Covance inventory if possible.
subject_df <- read.csv(dm_filepath)[c("USUBJID")] %>%
  mutate(PATNUM = str_extract(USUBJID, "[^-][^-]*$")) %>% distinct()

```

### Visits
```{r}

tv_file <- read_sas(tv_filepath) %>% mutate_if(is.character, str_trim) %>% mutate_if(is.character, toupper)

tv_VISIT <- unique(tv_file$VISIT) %>% toupper() %>% sort()
tv_VISIT

```

### ZB domain
```{r}

zb_file <- read_sas(zb_filepath) %>% 
  mutate_if(is.character, str_trim) %>%
  mutate_if(is.character, toupper) %>%
  # Replace empty values with NA
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), "NA", .))

# Vendors
zb_file$ZBNAM %>% unique()

# SAMPLE_ID columns 
(c("ZBREFID","ZBGRPID","ZBSPID")) %in% names(zb_file)

```

### TDQ

```{r}

# Collect information about tissue samples from TDQ file
tdq_df <- read.csv(tdq_path) %>%
  select(`STUDYID` = `PROJECT`,
         `PATNUM` = `SUBJECT`,
         `SITEID` = `SID`,
         `SAMPLE_ID` = `RREFID`,
         `Y` = `SAMP_DC_YYYY`,
         `M` = `SAMP_DC_MM`,
         `D` = `SAMP_DC_DD`,
         `BSETM` = `SAMP_TM`,
         `BMFLOC` = `LOCBIO_STD`,
         `PFSPEC` = `SAMPP_STD`) %>%
  mutate(BSED = paste(Y, M, D, sep = "-")) %>% distinct()

```

## Central lab(s)

#### Covance
Covance is the central lab.  
**
Note to curators:  
* Covance sample IDs are usually a concatenation of the Covance accession number and 2 digits reflecting the container number. Covance accession number usually has a format [A-Z]{1}[0-9]{6} (e.g. Z151423) or [0-9]{10} (e.g. 6500267332). The container number could be added with (e.g. Z151423-17, 6500267332-14) or without the dash (e.g. Z15142317, 650026733214) to make the full sample ID.  
**
```{r}

cova_inventory <- read_xls(path = cova_inventory_filepath, sheet = "Specimen Details", col_types = "text", na = "", skip = 1) %>%
  # Select and rename columns
  select(`STUDYID` = `Protocol`,
         `PATNUM` = `Patient Number`,
         `SCRNUM` = `Screening Number`,
         `ACCSNM` = `Accession Number`,
         `CONTAINER` = `Container Number`,
         `SAMPLE_ID` = `Container Barcode`,
         `BSED` = `Spec. Coll. Date`,
         `VISIT` = `Visit`,
         `PFSPEC` = `Specimen Type`,
         `SHIP_TO` = `Ship To`) %>%
  mutate_all(toupper) %>%
  mutate(BSED = format(as.Date(BSED, "%d-%b-%Y"),bsed_format),
         BMFNAM = "COVA",
         PRACCSNM = NA,
         DERIVED_FROM = SAMPLE_ID,
         IS_PRIMARY = "TRUE",
         BSETM = NA,
         # Harmonize PFSPEC values. Modify as necessary.
         PFSPEC = case_when(grepl("Slides", PFSPEC, ignore.case = T) ~ "SLIDE",
                            grepl("Blood", PFSPEC, ignore.case = T) ~ "BLOOD",
                            grepl("Serum", PFSPEC, ignore.case = T) ~ "SERUM",
                            grepl("Plasma", PFSPEC, ignore.case = T) ~ "PLASMA",
                            TRUE ~ PFSPEC),
         # Create BMFLOC. Modify as necessary.
         BMFLOC = case_when(grepl("PBMC|PLASMA|SERUM|BLOOD", PFSPEC) ~
                              "WHOLE BLOOD",
                            TRUE ~ "NA")) %>%
  # Remove potentially duplicated samples
  distinct(`SAMPLE_ID`, .keep_all = TRUE)

```
Covance inventory contains the SCRNUMs
```{r}
# Add SCRNUMs to subjects_df from Covance inventory
subject_df <- subject_df %>% 
  full_join(cova_inventory %>% select("PATNUM","SCRNUM") %>% distinct() %>%
              drop_na(PATNUM) %>% drop_na(SCRNUM), by="PATNUM")

```

## Assays

**
Note to curators.
Formatting convention:
### Assay name (= folder name in the study folder usually)

#### Vendor name  
* Brief description of the sample flow for the assay/vendor combination.

##### Rawdata  
* Brief comment if necessary.  
[Code chunk for processing raw data files]

##### Sample Metadata  
* Brief comment if necessary.  
[Code chunk for processing sample manifests/inventories etc.]  
The output should a dataframe formatted as a "mini output table". You should have at least all DSI columns and may include extra columns.
**

### Histology

#### QLAB

##### Rawdata

##### Sample Metadata

```{r}

# List all folders in histology folder.
# Modify pattern as needed.
ihc_qlab_folders <- list.files(path= hist_rawdata_path, pattern = "qlab",
                               full.names= TRUE, ignore.case = FALSE)

# List all manifest files in ihc_image_folders
# Modify pattern as needed.
ihc_qlab_files <- list.files(path=ihc_qlab_folders, pattern = ".csv",
                             full.names = TRUE, ignore.case = FALSE)
ihc_qlab_files

ihc_qlab_df <-CombineCSVFiles(ihc_qlab_files) %>% 
  select(`STUDYID` = `Study Identifier`,
         `BMFNAM` = `Vendor Name`,
         `SCRNUM` = `Unique Subject Identifier`,
         `VISIT` = `Visit Code`,
         `BSED` = `Date of Specimen Collection`,
         `BSETM` = `Time of Specimen Collection`,
         `SAMPLE_ID` = `Vendor Unique Sample Identification`,
         `BMFLOC` = `Anatomical Location of  Specimen`,
         `PFSPEC` = `Specimen Type`,
         `SPECCOND` = `Specimen Condition`) %>%
  mutate(BSED = format(as.Date(BSED, "%Y%m%d"), bsed_format),
         PFSPEC = case_when(grepl("BLOCK SLIDE", PFSPEC, ignore.case = T) ~
                              "BLOCK",
                            TRUE ~ NA_character_))

# Remove duplicate sample IDs
ihc_qlab_df <- ihc_qlab_df %>% distinct(SAMPLE_ID, .keep_all = TRUE) %>%
  filter(!is.na(SAMPLE_ID))

# Make sure all samples have a subject identifier
ihc_qlab_df[is.na(ihc_qlab_df$SCRNUM),]

# Add patient number and USUBJIDs
ihc_qlab_df <- ihc_qlab_df %>%
  left_join(subject_df) %>% distinct()

# Example code to solve duplicate ID issue
sample_occur <- as.data.frame(table(ihc_qlab_df$SAMPLE_ID))
ihc_qlab_df_dup <- ihc_qlab_df %>% 
  filter(SAMPLE_ID %in% sample_occur[sample_occur$Freq > 1, ]$Var1)
ihc_qlab_df <- ihc_qlab_df %>% anti_join(ihc_qlab_df_dup)

ihc_qlab_df_dup <- ihc_qlab_df_dup %>% filter(!is.na(PATNUM))
ihc_qlab_df <- ihc_qlab_df %>% bind_rows(ihc_qlab_df_dup)

# Add subject identifiers
ihc_qlab_df <- ihc_qlab_df %>% left_join(subject_df)

# Add lineage information
ihc_qlab_df$PRACCSNM <- NA
ihc_qlab_df$DERIVED_FROM <- ihc_qlab_df$SAMPLE_ID
ihc_qlab_df$IS_PRIMARY <- "TRUE"

ihc_qlab_df_final <- ihc_qlab_df[, output_sample_mapping_table_colnames]
rm(ihc_qlab_df)

# Example code for changing value manually
ihc_qlab_df_final[ihc_qlab_df_final$SAMPLE_ID=="AF461895",]$BSED<-"2014-03-12"

```

#### Histogenex
[...] sent slides and tissue blocks to Histogenex for DNA/RNA extraction. Histogenex then sent RNA to EA to perform RNAseq. When Histogenex extracted the RNA sample, each RNA sent to EA was assigned an RNA sample ID by concatenating a 5 letter-digit suffix to the HGX tissue ID. 
```{r}

# Common Histogenex (HIST) samples and their PFSPEC: The 9-digits HIST IDs are tissue IDs (e.g. 05511463C). The 14-digits HIST IDs are RNA, DNA or slide IDs (e.g. 05851623B0015R is RNA ID)
hist_id_pattern_short <- "[0-9]{8}[A-Z]"
hist_id_pattern_long <- paste0(hist_id_pattern_short, "[0-9]{4}[A-Z]")
hist_id_pattern <- paste0(hist_id_pattern_short, "|", hist_id_pattern_long)

```

##### Rawdata
Images
```{r}

# Extract sample IDs from names of images
ihc_hist_image_ids <- list.files(path = file.path(hist_rawdata_path, "image"),
                                 full.names = TRUE, recursive = TRUE) %>%
  toupper() %>% str_extract(hist_id_pattern_long) %>% unique()

length(ihc_hist_image_ids)

```

##### Sample Metadata

```{r}

# List all folders in histology folder.
# Modify pattern as needed.
ihc_hist_folders <- list.files(path= hist_rawdata_path, pattern = "hist_",
                               full.names= TRUE, ignore.case = FALSE)

# List all manifest files in ihc_image_folders.
# Modify pattern as needed.
ihc_hist_files <- list.files(path=ihc_hist_folders, pattern = ".csv",
                             full.names = TRUE, ignore.case = FALSE)
ihc_hist_files

ihc_hist_df <- CombineCSVFiles(ihc_hist_files) %>% 
  select(`STUDYID` = `Study Identifier`,
         `BMFNAM` = `Vendor Name`,
         `SCRNUM` = `Unique Subject Identifier`,
         `VISIT` = `Visit Code`,
         `BSED` = `Date of Specimen Collection`,
         `BSETM` = `Time of Specimen Collection`,
         `SAMPLE_ID` = `Vendor Unique Sample Identification`,
         `PFSPEC` = `Specimen Type`,
         `SPECCOND` = `Specimen Condition`) %>% 
  mutate(BSED = format(as.Date(BSED, "%Y%m%d"), bsed_format),
         PFSPEC = case_when(grepl("BLOCK SLIDE", PFSPEC, ignore.case = T) ~
                              "BLOCK",
                            TRUE ~ NA_character_))

# Remove duplicate rows based on Sample_ID
ihc_hist_df <- ihc_hist_df %>% distinct(SAMPLE_ID, .keep_all = TRUE) %>%
  filter(!is.na(SAMPLE_ID))

```

Add lineage information from QLAB parental samples
```{r}

# [file name] of the file to be processed below for reference
# Links QLAB, Histogenex sample identifiers and PATNUM/SCRNUM but some PATNUMs missing cause issues so just use SCRNUM
qlab_hist_tracker_df <- read_xlsx(path = qlab_hist_tracker_filepath,
                                  sheet = "Overview", col_types = "text",
                                  na = "#N/A") %>% 
  select(`HIST_ID` = `Sample #`,
         `QLAB_ID` = `Sample Id`,
         `SCRNUM` = `Screening ID`) %>%
  left_join(subject_df) %>%
  # Remove potentially duplicated entries based on HIST_ID
  distinct(HIST_ID, .keep_all = TRUE)  

# Example code to add raw manually to the data frame
qlab_hist_tracker_df <- qlab_hist_tracker_df %>% add_row(QLAB_ID = "EB553863",
                                                         SCRNUM = "26369901")

## Build lineage

## Add image identifiers
ihc_hist_image_df <- data.frame(IMAGE_ID = ihc_hist_image_ids) %>% 
  mutate(HIST_ID = str_extract(IMAGE_ID,hist_id_pattern_short)) %>%
  drop_na(IMAGE_ID) %>% distinct()

setdiff(ihc_hist_image_df$HIST_ID, ihc_hist_qlab_df$HIST_ID)
ihc_hist_qlab_df <- ihc_hist_qlab_df %>% left_join(ihc_hist_image_df)

## Format final dataframe

## Image level section
ihc_hist_df_images <- ihc_hist_qlab_df %>% 
  mutate(PRACCSNM = HIST_ID,
         SAMPLE_ID = IMAGE_ID,
         DERIVED_FROM = QLAB_ID,
         PFSPEC = "SECTION",
         BMFNAM = "HIST",
         IS_PRIMARY = "FALSE")

ihc_hist_df_images <- 
  ihc_hist_df_images[, sample_mapping_table_colnames] %>% distinct()

## Block or slide level
ihc_hist_df_sb <- ihc_hist_qlab_df %>% 
  mutate(PRACCSNM = QLAB_ID,
         SAMPLE_ID = HIST_ID,
         DERIVED_FROM = QLAB_ID,
         PFSPEC = case_when(grepl("[S]{1}$", HIST_ID) ~ "SECTION",
                            grepl("[X]{1}$", HIST_ID) ~ "SLIDE",
                            grepl("[B]{1}$", HIST_ID) ~ "BLOCK",
                            grepl("[P]{1}$", HIST_ID) ~ "BLOOD",
                            grepl("SLIDE CASE", HIST_ID, ignore.case = T) ~
                              "SLIDE",
                            TRUE ~ NA_character_),
         BMFNAM="HIST",
         IS_PRIMARY="FALSE")

# Put all the columns in the order for output table
ihc_hist_df_sb <- ihc_hist_df_sb[, output_sample_mapping_table_colnames] %>%
  distinct()

## Check we already have all QLAB parents
setdiff(ihc_hist_qlab_df$QLAB_ID, ihc_qlab_df_final$SAMPLE_ID)

## Compare metadata
compare_tables(ihc_hist_qlab_df %>%
                 select(SCRNUM, VISIT, BSED, BSETM, QLAB_ID, USUBJID, PATNUM) %>%
                 distinct(),
               ihc_qlab_df_final %>% select(lineage_inherited_colnames) %>%
                 rename(`QLAB_ID` = `SAMPLE_ID`) %>% distinct(),
               key = c("QLAB_ID"))

## Combine all samples for IHC assay
ihc_hist_df_final <- ihc_hist_df_images %>%
  rbind(ihc_hist_df_sb) %>% drop_na(SAMPLE_ID)

rm(ihc_hist_df)

```


### NGS

#### DNAseq FDMI

##### Rawdata
Read the FMI "one-sample-per-line" raw data files.
```{r dnaseq ngs, message=FALSE, warning=FALSE}

# Function to extract character date and convert it to actual date
extract_date <- function(x){
  if(!is.na(x)){
    return(case_when(
      str_detect(x, "20\\d{6}") ~ 
        as.Date(str_extract(x, "20\\d{6}"), format = "%Y%m%d"),
      str_detect(x, "\\d{2}[A-z]{3}20\\d{2}") ~
        as.Date(str_extract(x, "\\d{2}[A-z]{3}20\\d{2}"), format = "%d%b%Y")
    )) 
  }
}

# Make a vector of all the "one-sample-per-line" files.
# Note: not subsetting using "std" since sometimes it is absent in filenames
fmi_data_files <- ngs_rawdata_path %>% dir(full.names=T, recursive=T) %>% 
  str_subset("one-sample-per-line") %>% str_subset("xlsx|XLSX")

# Assign the key word and the date to each FMI "one-sample-per-line" file.
# E.g. in file name
# ngs_dna_targdna_foundationone_15-143_n154_one-sample-per-line_20160113.xlsx
# the key word is 15-143.
# Note: key word is unique to each FMI experiment.

fmi_data_files_df <- data.frame(File_Path = fmi_data_files,
                                stringsAsFactors = F) %>%
  mutate(File_Name = basename(File_Path)) %>%
  mutate(FMI_key_word = str_extract(File_Name, "\\d{2,}-\\d{2,}"),
         # extract_date function is wrapped in suppressWarnings as 
         # all the statements in the case_when function are evaluated every time,
         # so it generates a warning every time when one of the
         File_Date = suppressWarnings(extract_date(File_Name))) %>% 
  group_by(FMI_key_word) %>% mutate(Latest_File_Date = max(File_Date))

# Find the latest "one-sample-per-line" file version for a given FMI keyword
fmi_data_files_2process_df <- fmi_data_files_df %>%
  filter(File_Date == Latest_File_Date)

print("These are the latest one-sample-per-line FMI manifests that will be processed:")
print.data.frame(fmi_data_files_2process_df[, c("File_Name", "File_Date")])

fmi_data_files_2process <- as.vector(fmi_data_files_2process_df$File_Path)

# Function for stitching together data files in xlsx format
# Column BAITSET is read to separate assays that use tissue samples
# from those that use liquid samples (when applicable).
CombineDataFiles <- function(file_list){
  df <-  data.frame(stringsAsFactors=FALSE)
  for(i in seq_along(file_list)){
    tryCatch({xlsx_file <- read_xlsx(file_list[i], col_names = TRUE,
                                     col_types = "text", na = "-")},
             error=function(e) {cat("Bad file pattern:\n", file_list[i], "\n")}
    )
    # File that doesn't have the column "BAITSET" (older format):
    if("BAITSET" %notin% colnames(xlsx_file)) {
      j <- xlsx_file %>% select(`SAMPLE_ID` = `FMI SAMPLE ID`,
                                `PATNUM` = `SUBJECT ID`,
                                `PRACCSNM` = `SAMPLE ID`) %>%
        mutate(BAITSET = NA_character_)
      df <- bind_rows(df, j)
      # File that has the column "BAITSET" (current format):
    } else {
      j <- xlsx_file %>% select(`SAMPLE_ID` = `FMI SAMPLE ID`,
                                `PATNUM` = `SUBJECT ID`,
                                `PRACCSNM` = `SAMPLE ID`,
                                `BAITSET`)
      df <- bind_rows(df, j)
    }
  }
  return (df)
}

fmi_raw_df <- CombineDataFiles(fmi_data_files_2process) %>%
  distinct(SAMPLE_ID, .keep_all = TRUE) %>%
  filter(!is.na(SAMPLE_ID))

```

##### Sample Metadata
Add the metadata inherited from parental HIST samples.
```{r}

# Check whether all the parental samples are present
setdiff(fmi_raw_df$PRACCSNM, ihc_hist_df_final$SAMPLE_ID)

# Add the metadata from parental HIST samples and apply output table format
ngs_fdmi_df_final<-fmi_raw_df %>%
  left_join(ihc_hist_df_final %>%
              select(lineage_inherited_colnames) %>%
              rename("PRACCSNM" = "SAMPLE_ID") %>% distinct()) %>%
  mutate(PFSPEC = "DNA",
         BMFNAM = "FDMI",
         IS_PRIMARY = "FALSE") %>%
  select(output_sample_mapping_table_colnames) %>% distinct()

```

#### RNAseq EA(Q2SO)
EA has its own specific sequencing IDs in their LIMS. However, the EA IDs are not usually used by EA when transferring sequencing data to file share. RNA sample IDs are typically used instead. 

**
Note to curators:  
* If Histogenex sent RNA sample to EA, typically 2 "layers" of samples need to be added to the output table:  
1. "layer" of samples with IDs like 07290002C0017R with BMFNAM = HIST, PFSPEC = RNA to reflect that Histogenex extracted RNA (see section "Histogenex" of this Rmd for more information about typical Histogenex sample IDs),  
2. "layer" of samples with the same IDs with BMFNAM = Q2SO, PFSPEC = RNA to reflect that Q2SO (EA) performed NGS on these RNA samples.  
**

##### Rawdata
``` {r, rnaseq ngs, message=FALSE, warning=FALSE}

# Check raw data files and compare with the manifest to see whether any samples are missing

# List all NGS assay folders
ngs_folders <- list.files(path= ngs_rawdata_path,
                          pattern = "ngs_rna", full.names= TRUE,
                          ignore.case = FALSE)

# List all qc_summary files
qc_summary_files <- list.files(path=ngs_folders, pattern = "qc_summary",
                               full.names = TRUE, ignore.case = FALSE)

# Function to combine all qc_summary files
ea_combined <- function(file){
  ea_df <- data.frame(EA_ID=character(), stringsAsFactors=FALSE)
  for(i in file){
    j = as.data.frame(read.delim(i, header = TRUE, sep = "\t", fill = TRUE, stringsAsFactors = FALSE)[2])
    colnames(j) <- "SAMPLE_ID"
    ea_df <- rbind(ea_df, j)
  }
  ea_df
}

ea_raw_df <- ea_combined(qc_summary_files) %>% distinct() %>%
  mutate(HIST_ID = str_extract(SAMPLE_ID, hist_id_pattern_short))

# Samples containing "UHRR" (Universal Human Reference RNA) in their names are controls that don't need to be mapped
ea_raw_df[is.na(ea_raw_df$HIST_ID),]$SAMPLE_ID
# Remove controls
ea_raw_df <- ea_raw_df %>% drop_na(HIST_ID)
```

##### Sample Metadata
Add the metadata inherited from parental HIST samples.
```{r}

## RNAseq assay -> PFSPEC = "RNA"
ngs_ea_df <- ea_raw_df %>% 
  mutate(BMFNAM = "Q2SO",
         PFSPEC = "RNA",
         IS_PRIMARY = "FALSE",
         PRACCSNM = SAMPLE_ID)

## Add lineage
setdiff(ea_raw_df$HIST_ID, ihc_hist_df_final$SAMPLE_ID)
# final df 

# With Q2SO samples
ngs_ea_df <- ngs_ea_df[!ngs_ea_df$HIST_ID %in% setdiff(ea_raw_df$HIST_ID,ihc_hist_df_final$SAMPLE_ID),] %>%
  left_join(ihc_hist_df_final %>% select(lineage_inherited_colnames) %>%
              rename(HIST_ID="SAMPLE_ID")) 

# Add RNA by HIST
ngs_ea_df_final<-ngs_ea_df %>% rbind(ngs_ea_df %>%
                                       mutate(BMFNAM="HIST",PRACCSNM=HIST_ID)) %>%
  select(output_sample_mapping_table_colnames) %>% distinct()

rm(ea_raw_df)
rm(ngs_ea_df)

```

##### NGS pipeline IDs
NGS pipeline IDs are generated during NGS data analysis process. They are child samples of NGS samples and should be included in output table if their sample IDs are different from those of NGS samples.
```{r ngs pipeline, results='hide', message=FALSE, warning=FALSE}

pipeline_path <- file.path(ngs_outdata_path, "ngs_pipeline")
pipeline_folders <- list.files(pipeline_path, pattern = "ea_obd",
                               full.names = TRUE)
pipeline_folders <- pipeline_folders %>% paste0("/configs")
pipeline_files <- list.files(pipeline_folders, pattern = ".csv",
                             full.names = TRUE)

# Function to combine all pipeline files
pipeline_combined <- function(file){
  df <-  data.frame(Characters=character(), Characters=character(), stringsAsFactors=FALSE)
  for(i in file){
    j = as.data.frame(read_csv(file=i, col_types = cols(.default = col_character())))["SAMID"]%>%mutate(SAMID=toupper(SAMID))
    colnames(j) <- "Pipeline_ID"
    df <- rbind(df, j)
  }
  df
}

pipeline_df <- pipeline_combined(pipeline_files)
pipeline_df <- pipeline_df %>% distinct()

setdiff(pipeline_df$Pipeline_ID, ngs_ea_df_final$SAMPLE_ID)

#add to final assay df
#ngs_ea_df_final <- ngs_ea_df_final %>% rbind(
#  ngs_ea_df_final %>%filter(BMFNAM=="Q2SO") %>% right_join(pipeline_df%>%rename(SAMPLE_ID=Pipeline_ID)) %>% mutate(BMFNAM="GNE")
#)
```

# Merge output table
```{r output table, results='hide', message=FALSE, warning=FALSE}

output_table <- ihc_hist_df_final %>%
  bind_rows(ihc_qlab_df_final) %>%
  bind_rows(ngs_fdmi_df_final) %>%
  bind_rows(ngs_ea_df_final)

# Convert everything to upper case
output_table <- mutate_all(output_table, .funs=toupper)

output_table <- output_table %>%
  distinct() %>%
  arrange(SCRNUM)

```

# Samples from zb
**
Note to curators:  
* Columns in zb file that can have sample ids: ZBGRPID, ZBREFID, ZBSMPLID. If there are sample IDs in these columns, check if those sample IDs are already in the output table.  
* If a given sample ID is found in both zb file and output table, check if the values from other columns in zb (STUDYID, USUBJID, ZBNAM, ZBSPEC, VISIT, ZBDTC) match corresponding values in the output table. If not, try to find out from BOM which values (from zb or from output table) are correct.  
* If a certain sample ID is missing from the output table, add it to the output table along with other information about this sample available (STUDYID, USUBJID, ZBNAM, ZBSPEC, VISIT, ZBDTC). Deduce PATNUM and SCRNUM values for this sample based on USUBJID if possible. Build sample lineage and assign values to PRACCSNM, DERIVED_FROM and IS_PRIMARY columns if possible. Assign NA to any columns for which you cannot find the values (like BSETM, BMFLOC).  
**

```{r}

zb_add <- zb_file %>% 
  select(`STUDYID`,
         `USUBJID`,
         `QLAB_ID` = `ZBREFID`,
         `BMFNAM` = `ZBNAM`,
         `PFSPEC` = `ZBSPEC`,
         `SPECCOND` =`ZBSPCCND`,
         `BMFLOC` =`ZBLOC`,
         `VISIT`,
         `HIST_ID`=`ZBSMPLID`) %>% distinct()

setdiff(zb_add$QLAB_ID, output_table$SAMPLE_ID)
setdiff(zb_add$HIST_ID, output_table$SAMPLE_ID)

compare_tables(zb_add %>% rename(SAMPLE_ID = QLAB_ID),
               output_table %>% filter(BMFNAM == "QLAB"),
               key = c("SAMPLE_ID"))

compare_tables(zb_add %>% rename(SAMPLE_ID = HIST_ID),
               output_table %>% filter(BMFNAM == "HIST"),
               key=c("SAMPLE_ID"))

```


# Harmonization

## Subjects
In the rawdata there were additional SCRNUM-PATNUM pairs. "Propagate" the found values.
```{r}

subj_dsi <- output_table %>% select(PATNUM, SCRNUM) %>% distinct() %>% drop_na(PATNUM)

count(subj_dsi, SCRNUM) %>% filter(n > 1)
count(subj_dsi, PATNUM) %>% filter(n > 1)

#Example code to manually add SCRNUM
subj_dsi[subj_dsi$PATNUM == "406103",]$SCRNUM <- "26688806"

#save and use when rerun script
#write.csv(subj_dsi %>% distinct(), scrnum_patnum_filepath, row.names = FALSE)
```

## Metadata

```{r}

# Use as needed

# Harmonize BSED
output_table$BSED <- ifelse(!is.na(output_table$BSED),
                        format(strptime(output_table$BSED, "%Y%m%d"),
                               format = "%Y-%m-%d"), NA)

# Harmonize BSETM
output_table$BSETM <- ifelse(!is.na(output_table$BSETM),
                          format(strptime(output_table$BSETM, "%I:%M"),
                                 format="%H:%M"), NA)
# Change 00:00, 00:01, 23:59 to NA
output_table$BSETM <- ifelse(output_table$BSETM %in% c("00:00","00:01","23:59"), NA,
                          output_table$BSETM)


# Harmonize BMFLOC
output_table <- output_table %>%
  mutate(BMFLOC = case_when(grepl("OTHER; SPECIFY BELOW:", BMFLOC) ~
                              gsub("OTHER; SPECIFY BELOW:", "", BMFLOC),
                            grepl("OTHER:", BMFLOC) ~
                              gsub("OTHER:", "", BMFLOC),
                            grepl("NOT SPECIFIED ON REPORT", BMFLOC) ~ NA,
                            grepl("UNKNOWN EXACT LOCATION", BMFLOC) ~ NA,
                            grepl("UNKNOWN", BMFLOC) ~ NA,
                            grepl("UNK", BMFLOC) ~ NA,
                            TRUE ~ BMFLOC))

# Harmonize PFSPEC
output_table <- output_table %>%
  mutate(PFSPEC = case_when(grepl("[^BLOCK]SLIDE", BMFLOC) ~ "SLIDE",
                            grepl("BLOCK", BMFLOC) ~ "BLOCK",
                            # "WHOLE BLOOD" -> "BLOOD"
                            grepl("BLOOD", PFSPEC) ~ "BLOOD",
                            grepl("RNA", PFSPEC) ~ "RNA",
                            grepl("DNA", PFSPEC) ~ "DNA",
                            TRUE ~ PFSPEC))

# Harmonize PFSPEC of Histogenex samples based on identifiers
output_table <- output_table %>%
  mutate(PFSPEC = case_when(BMFNAM == "HIST" & grepl("B$", SAMPLE_ID) ~ "BLOCK",
                            BMFNAM == "HIST" & grepl("C$", SAMPLE_ID) ~ "SLIDE",
                            BMFNAM == "HIST" & grepl("S$", SAMPLE_ID) ~ "SECTION",
                            BMFNAM == "HIST" & grepl("R$", SAMPLE_ID) ~ "RNA"
                            BMFNAM == "HIST" & grepl("D$", SAMPLE_ID) ~ "DNA"
                            BMFNAM == "HIST" & grepl("P$", SAMPLE_ID) ~ "BLOOD"
                            TRUE ~ PFSPEC))


# Harmonize VISIT
output_table <- output_table %>%
  mutate(VISIT = case_when(grepl("^C[0-9]", VISIT) ~
                             str_replace(VISIT, "^C", "CYCLE "),
                           grepl("D[0-9]", VISIT) ~
                             str_replace(VISIT, "D", " DAY "),
                           grepl("UNSCH", VISIT) ~
                             "UNSCHEDULED",
                           grepl("SCRN", VISIT) ~ "SCREENING",
                           grepl("FRESH BIOPSY", VISIT) ~ NA,
                           TRUE ~ VISIT))

# Remove *
output_table <- output_table %>% 
    mutate_all(funs(str_replace_all(., "\\*", "")))

# Use tv file
tv_file <- read_sas(tv_filepath)
tv_VISIT <- unique(tv_file$VISIT) %>% toupper() %>% sort()
# VISIT code that need to be harmonized
setdiff(output_table$VISIT %>% unique(), tv_VISIT)

# Replacement might have introduced empty string, replace empty string by NA
output_table<-output_table %>% distinct() %>% 
  mutate_all(list(~ na_if(., "")))

output_table[output_table == "NA"] <- NA
output_table[output_table == " "] <- NA

```

### PFSPEC, SPECCOND, SPECFRMT
This section is for upcoming harmonization, not used yet.
```{r, eval=FALSE}


```

### Final Formatting

```{r QC issues and final format, message=FALSE, warning=FALSE}

# Use as needed

# Remove commas which causes issues for csv format
output_table <- output_table %>% 
    mutate_all(funs(str_replace_all(., ",", " ")))

# Change all entries to upper case
output_table <- mutate_all(output_table, .funs=toupper)

# Trim all white space
output_table <- mutate_all(output_table, list(~ trimws(.)))

# Distinct, no NA sample id
output_table <- distinct(output_table, .keep_all = TRUE) %>% 
  mutate_all(~ ifelse(. %in% c("N/A", "null", ""), "NA", .)) %>% 
  drop_na(SAMPLE_ID)

output_table$STUDYID %>% unique()
output_table$STUDYID <- study %>% toupper()

```

# Save the output table
```{r}

output_table <- output_table[, output_sample_mapping_table_colnames] %>% 
  drop_na("SAMPLE_ID") %>%
  distinct() %>% 
  arrange(SCRNUM)

write.csv(output_table, na = "NA", row.names = FALSE, 
          file = output_table_path, quote = FALSE)

```

# R session info
This document was created with the following R session parameters:
```{r session info}

session_info()

```
